---
title: "Homework 2 Technical Document"
author: "Group 6"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(caret)
data <- read_csv("XYZData.csv")
```

# Table of contents:
1. Problem Scoping
2. Exploratory Data Analysis
3. Predictive Analysis
4. Summary for Analysis Results

# 1. Problem Scoping
## 1.1 Background Information
Website XYZ, a music-listening social networking website, follows the "freemium" business model. The website offers basic services for free, and provides a number of additional premium capabilities for a monthly subscription fee.

## 1.2 Problem Statement
Know which users would be likely to convert from free users to premium subscribers in the next 6 month period, if they are targeted by the promotional campaign.

## 1.2 Objective
Build the best predictive model to predict likely adopters for the next marketing campaign

## 1.3 Data Requirement
The user features and results of previous marketing campaigns which targeted a number of non-subscribers

## 1.4 Deliverables
1. A proper performance metric for model evaluation based on the description of the business
2. Sampling techniques to process imbalanced class distribution data
3. The best model that achieves highest performance on the metric of selection after comparing different modeling techniques, different parameter configurations, and feature selection strategies.

# 2. Exploratory Data Analysis
## 2.1 Get a summary of dataset
```{r}
summary(data |> select(-user_id))
```
Most of features are positively skewed except for male, avg_friend_male, delta_avg_friend_male, delta_subscriber_friend_cnt, and tenure.

## 2.2 Missing value check
```{r}
sum(is.na(data))
```
There is no row with missing values in the dataset.

## 2.3 Visualize the adopter
```{r}
# pie chart for adopter to see the proportion
ggplot(data, aes(x = factor(adopter), fill = factor(adopter))) +
  geom_bar()
```
```{r}
num_class1 <- data |> filter(adopter==1) |> count()
as.numeric(100*num_class1/count(data))  # in percentage
```
The class distribution is highly imbalanced with only 3.7% are adopters (class 1).

## 2.4 Explore relationships between variables
```{r}
# Create a correlation matrix
cor_matrix <- round(cor(data[,2:27]),2)
# Find the top 5 correlated column pairs
top_cor <- findCorrelation(cor_matrix, cutoff = 0.5, names = TRUE)
# Print the top correlated column pairs
print(top_cor)
```
These features have strong correlation (correlation coefficient >= 0.5):

1. age-related
  + cor(age, avg_friend_age) = 0.69
2. friend-related
  + cor(friend_cnt, friend_country_cnt) = 0.69
  + cor(friend_cnt, subscriber_friend_cnt) = 0.65
  + cor(delta_friend_cnt, delta_friend_country_cnt) = 0.65
  + cor(friend_country_cnt, subscriber_friend_cnt) = 0.57
3. track-related
  + cor(lovedTracks, delta_lovedTracks) = 0.51

# 3. Predictive Analysis
## 3.1 Performance metric
```{r}

```
## 3.2 Oversampling on training data
```{r}

```
## 3.3 Model1
```{r}

```
## 3.4 Model2
```{r}

```
## 3.5 Performance Comparision

# 4. Summary for Analysis Results
